Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to boundary detection tasks

### 2023-08-12 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits that enduring, high-level musical elements persist regardless of sonic changes and constantly evolve within their context, mirroring the nature of sheet music's symbolic domain. A single waveform can correspond to a finite set of representations. Still, one music piece ---or music representation--- can yield infinite interpretations, varying by factors such as instruments and performance styles, to name just a few.

@@

For example, despite differing cultural contexts, styles, time and key signatures, and sonic properties, we suggest a commonality between Figure \ref{fig:mahler} and Figure \ref{fig:giant_steps}. Both pieces display the same 'musicological flavor', derived from their melodic and harmonic contour envelopes through non-diatonic major thirds.

This concept aligns with the theories of Kurt Koffka, a founding member of the Gestalt school of psychology, who advocated for a holistic approach to understanding complex forms, as opposed to the structuralist practice of dissecting mental processes into their elemental components \cite{Koffka2013PrinciplesPsychology}. In the context of music, Gestalt psychology principles suggest our minds process auditory input similarly to visual input by seeking patterns and structures. This understanding significantly enhances the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

\begin{figure}[ht]
    \centering
    \scalebox{0.9}{
    \includegraphics[width=\textwidth]{figures/images/Mahler 9 Giant Steps score.png}
        }
    \caption[Mahler's 9th Symphony, 2nd movement]{\small{A small excerpt from Mahler's 9th Symphony, 2nd movement: The melodic and harmonic contour propels through non-diatonic major thirds.}}
    \label{fig:mahler}
\end{figure}

\begin{figure}[ht]
    \centering
    \scalebox{0.9}{
    \includegraphics[width=\textwidth]{figures/images/giant steps score.png}
        }
    \caption[Giant Steps]{\small{John Coltrane's Giant Steps head, featuring rapid chord changes in non-diatonic major thirds.}}
    \label{fig:giant_steps}
\end{figure}

>>>>

This thesis posits that enduring, high-level musical elements persist regardless of sonic changes, mirroring the nature of the symbolic domain --sheet music--. A single waveform can correspond to a finite set of representations. Still, one sheet music excerpt can yield literally infinite interpretations.

@@

For example, despite differing cultural contexts, styles, time and key signatures, and sonic properties, we suggest a concept commonality between Figure \ref{fig:mahler} and Figure \ref{fig:giant_steps}. Both pieces display the same 'musicological flavor' derived from their melodic and harmonic contour envelopes through non-diatonic major thirds.

This concept aligns with the theories of Kurt Koffka, a founding member of the Gestalt school of psychology, who advocated for a holistic approach to understanding complex forms, as opposed to the structuralist practice of dissecting mental processes into their elemental components \cite{Koffka2013PrinciplesPsychology}. In the context of music, Gestalt psychology principles suggest that our minds process auditory input similarly to visual input by seeking patterns and structures. This understanding significantly enhances the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

\begin{figure}[ht]
    \centering
    \scalebox{0.9}{
    \includegraphics[width=\textwidth]{figures/images/Mahler 9 Giant Steps score.png}
        }
    \caption[Mahler's 9th Symphony, 2nd movement excerpt.]{\small{A small excerpt from Mahler's 9th Symphony, 2nd movement: The melodic and harmonic contour propels through non-diatonic major thirds.}}
    \label{fig:mahler}
\end{figure}

\begin{figure}[ht]
    \centering
    \scalebox{0.9}{
    \includegraphics[width=\textwidth]{figures/images/giant steps score.png}
        }
    \caption[Giant Steps excerpt.]{\small{John Coltrane's Giant Steps head, featuring rapid chord changes in non-diatonic major thirds.}}
    \label{fig:giant_steps}
\end{figure}

### 2023-08-12 introduction/Subsections/Introduction.tex
Updates from Overleaf

Prior research emphasizes the necessity of selecting suitable audio features to distinguish and label unique music segments for music segmentation tasks. The challenge lies in acquiring annotated data for these feature transformations, which can be time-consuming and expensive. In response, scholars have turned to unsupervised deep learning using readily available music audio \cite{deepfeaturesegment, Grill2015MusicAnnotations}. This method has substantially improved the generalization capacity of machine-listening models in downstream applications. Among others, segmentation algorithms have seen significant enhancements and delivered state-of-the-art results \cite{Hernandez-Olivan2021MusicFeatures, Li2023MERT:Training}.

@@

All in all, given the proven effectiveness of deep audio embeddings in existing research and their potential for transfer learning, they hold considerable promise; therefore, we will leverage deep convolutional neural net-
works to replace traditional
acoustic features with sound-agnostic
and content-sensitive embeddings with applications to boundary detection and tracking segmentation tasks

>>>>

Prior research emphasizes the necessity of selecting suitable audio features to distinguish and label unique music segments for music segmentation tasks. The challenge lies in acquiring annotated data for these feature transformations, which can be time-consuming and expensive. In response, scholars have turned to unsupervised deep learning using readily available audio data \cite{deepfeaturesegment, Grill2015MusicAnnotations}. This method has substantially improved the generalization capacity of machine-listening models in downstream applications. Among others, segmentation algorithms have seen significant enhancements and delivered state-of-the-art results \cite{Hernandez-Olivan2021MusicFeatures, Li2023MERT:Training}.

@@

All in all, given the proven effectiveness of deep audio embeddings in existing research and their potential for transfer learning, they hold considerable promise; therefore, we will leverage deep convolutional neural networks to replace traditional acoustic features with sound-agnostic and content-sensitive embeddings with applications for boundary detection and track segmentation tasks.

### 2023-08-12 introduction/Subsections/Motivation.tex
Updates from Overleaf

ES, a Swedish company, curates a vast global library of over 40,000 royalty-free music tracks and 90,000+ sound effects. Based in Barcelona, MTG specializes in innovative sound and music technologies, such as information retrieval, digital signal processing, interactive music systems, and computational musicology.

@@

Advancements in AI research, building on the theoretical groundwork laid by pioneers such as Alan Turing and Claude Shannon, have led to significant breakthroughs \cite{Vaswani2017AttentionNeed}. With AI being applied broadly in skyrocketing popular products \cite{OpenAI2023GPT-4Report}, researchers and corporations must stay abreast of this rapidly evolving field. The emergence of self-supervised models capable of learning embedding spaces to retrieve musical content from audio signals presents new opportunities. These models, which autonomously extract information from audio data, can potentially transform multiple aspects of the music industry. Industrially, these embedding spaces can be used to devise innovative products, enhance user experiences for content creators, and stimulate innovation and collaboration across the industry.

Traditionally, MIR research has focused on handcrafted features and annotated data. From an academic perspective, self-supervised models are worth exploring as they don't require annotated data and have demonstrated competitive performance in other tasks \cite{Li2023MERT:Training}.

>>>>

ES, a Swedish company, curates a vast global library of over 40,000 royalty-free music tracks and 90,000+ sound effects. Based in Barcelona, the MTG specializes in innovative sound and music technologies, such as information retrieval, digital signal processing, interactive music systems, and computational musicology.

@@

Advancements in AI research, building on the theoretical groundwork laid by pioneers such as Alan Turing and Claude Shannon, have led to significant breakthroughs \cite{Vaswani2017AttentionNeed}. With AI being applied broadly in skyrocketing popular tech products \cite{OpenAI2023GPT-4Report}, researchers and corporations must stay abreast of this rapidly evolving field.

The emergence of self-supervised models capable of learning embedding spaces to retrieve musical content from audio signals presents new opportunities. These models, which autonomously extract information from audio data, can potentially transform multiple aspects of the music industry. Industrially, these embedding spaces can be used to devise innovative products, enhance user experiences for content creators, and stimulate innovation and collaboration across the industry.

### 2023-08-12 methods/subsections/implementation.tex
Updates from Overleaf

\begin{figure}
    \centering
    \includegraphics{figures/images/applsci-08-00150-g003.png}
    \caption[SampleCNN filters]{\small{The spectrum of the estimated filters in the intermediate layers of SampleCNN is sorted by the frequency of the peak magnitude. The x-axis represents the index of the filter, and the y-axis represents the frequency range from 0 to 11 kHz. The model used for visualization is  39---SampleCNN with 59,049 samples as input. Visualization was performed using the gradient ascent method to obtain the accumulated gradient-based input waveform-like signal that maximizes the activation of a filter in the layers. To effectively find the filter characteristics, we set the input size to 729 samples, close to a typical frame size.}}
    \label{fig:filters}
\end{figure}

@@

\begin{table}[ht]
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash\hsize=0.4\hsize}X
  >{\centering\arraybackslash\hsize=0.2\hsize}X
  >{\centering\arraybackslash\hsize=0.2\hsize}X
  >{\centering\arraybackslash\hsize=0.2\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} & \thead{\textbf{Num. of layers}} \\
\midrule
Conv + ReLU & 1 & 128 & 1 \\
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 & 2 \\
 & 128 & 256 & 1 \\
 & 256 & 256 & 6 \\
 & 256 & 512 & 1 \\
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 & 1 \\
Fully connected layer & 512 & 128 & 1 \\
\bottomrule
\end{tabularx}
\caption[SampleCNN layer specifications]{\small{Layer specifications for our SampleCNN model implementation; extended from \cite{CLMR2021}.}}
\label{tab:samplecnntable}
\end{table}

>>>>

\begin{figure}
    \centering
    \includegraphics{figures/images/applsci-08-00150-g003.png}
    \caption[SampleCNN filters]{\small{The spectrum of the estimated filters in the intermediate layers of SampleCNN is sorted by the frequency of the peak magnitude. The x-axis represents the index of the filter, and the y-axis represents the frequency range from 0 to 11 kHz. The model used for visualization is  39---SampleCNN with 59,049 samples as input. \cite{Lee2018SampleCNN:Classification}}}
    \label{fig:filters}
\end{figure}

@@

\begin{table}[ht]
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash\hsize=0.4\hsize}X
  >{\centering\arraybackslash\hsize=0.2\hsize}X
  >{\centering\arraybackslash\hsize=0.2\hsize}X
  >{\centering\arraybackslash\hsize=0.2\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} & \thead{\textbf{Num. of layers}} \\
\midrule
Conv + ReLU & 1 & 128 & 1 \\
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 & 2 \\
 & 128 & 256 & 1 \\
 & 256 & 256 & 6 \\
 & 256 & 512 & 1 \\
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 & 1 \\
Fully connected layer & 512 & 128 & 1 \\
\bottomrule
\end{tabularx}
\caption[SampleCNN layer specifications]{\small{Layer specifications for our SampleCNN model implementation; extracted and extended from \cite{CLMR2021}.}}
\label{tab:samplecnntable}
\end{table}
